Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-07 doc/report/msc-report.tex
zohmg/process.py: fixed 'zohmg.util' import error.

From a user's perspective, storing aggregates in Zohmg is as simple as writing
a map (in the MapReduce sense) and letting Zohmg run that map over the data
set. The user's map is called 'usermap' from here on. In order to let Zohmg
know what output to expect from the mapper, the user also defines the
dimensions and units of his data. After all, the user most likely knows his
own data better than anyone else.

@@

Going back to the running example of apache query logs, a usermap might look
something like this:

\begin{verbatim}
# assumes Apache common log format.
def map(key, value):
  apachelog = parse_apache_log(value)
  dimensions = {'useragent' : apachelog.useragent,
                'path'      : apachelog.path,
                'status'    : apachelog.statuscode}
  values = {'pageviews' : 1,
            'bytes'     : apachelog.size}

@@

The actual parsing of the log is abstracted; it is the emitting of
dimensionality data I want to showcase in this example. We assume that
parse\_apache\_log() accepts a single line from the log file and returns an
object with the appropriate fields, such as useragent ant statuscode.

@@

The last line 'yields' the tuple. This is a python construct called a
generator that lets the map emit any number of tuples (or none!) for a single
invocation.

@@

This will instruct zohmg to apply the map function found in apache.py to all
logs that it can find in the /data/weblogs/2009/05/10 directory on HDFS. The
emitted output is duly collected and projections are carefully stored in the
data store.

@@

The cardinality of a dimension is the number of distinct values of the
dimension. For example, the dimension of 'logged in' has a cardinality of two
since 'logged in' is a boolean variable; the user is either logged in or not.
The 'country' dimension has a much higher cardinality: there are some 300
countries in the world. Certainly, there could be dimension whose cardinality
is infinite: the 'path' dimension in the web log example, for example, could
in theory have a limitless number of values.
The number of possible combinations of dimensions of your data -- the number
of points in n-space -- is directly related to the cardinality of each
dimension. It is likely that your data is of such a nature that it is
physically impossible to store information about all dimension combinations.
This is why we ask you, the user, to define what projections you are
interested in: storing aggregates for the slices of the cube that interest you
is both a time and space save, and in many cases the only feasible solution.

@@

[What are the limits of our system? How high cardinality do we support? (This
is closely tied to hbase and the qualifier-limit.) If Martin is going to have
'normalized_path' as one of his dimensions, we might run into trouble.]

@@

Dumbo expects a key-value pair from the mapper. Hadoop will collect all keys
and give the reducer a key and an iterator of values. If we are to store
projections, we need to encode both the desired dimensions and the unit in
question in the key and emit the value of the unit in the 'value' field. This
way we can use a simple SumReducer to sum the values for each unit at any
point in n-space and be done.

>>>>

From a user's perspective, storing aggregates in Zohmg is as simple as writing
a map (in the MapReduce sense) and letting Zohmg run that map over the data
set. The user's map is called \textit{usermap} from here on. In order to let Zohmg
know what output to expect from the mapper, the user also defines the
dimensions and units of his data. After all, the user most likely knows his
own data better than anyone else.

@@

Going back to the running example of apache access logs, a usermap might look
something like this:

\begin{verbatim}
def map(key, value):
  log = parse_apache_log(value)
  dimensions = {'useragent' : log.useragent,
                'path'      : log.path,
                'status'    : log.statuscode}
  values = {'pageviews' : 1,
            'bytes'     : log.size}

@@

The actual parsing of the log is abstracted; it is the emitting of
dimensionality data I want to showcase in this example. We assume that
\texttt{parse\_apache\_log()} accepts a single line from the log file and
returns an object with the appropriate fields, such as useragent ant
statuscode.

@@

The last line \textit{yields} the tuple. This is a python construct called a
generator that lets the map emit any number of tuples (or none!) for a single
invocation.

@@

This will instruct zohmg to apply the map function found in \texttt{apache.py}
to all logs that it can find in the \texttt{/data/weblogs/2009/05/10} directory
on HDFS. The emitted output is duly collected and projections are carefully
stored in the data store.

@@

The cardinality of a dimension is the number of distinct values of the
dimension. For example, the dimension of \textit{logged in} has a cardinality
of two since \textit{logged in} is a boolean variable; the user is either
logged in or not. The \textit{country} dimension has a much higher
cardinality: there are some 300 countries in the world. Certainly, there could
be dimension whose cardinality is infinite: the \textit{path} dimension in the
web log example, for example, could in theory have a limitless number of
values.

The number of possible combinations of dimensions of your data---the number
of points in n-space---is directly related to the cardinality of each
dimension. It is likely that your data is of such a nature that it is
physically impossible to store information about all dimension combinations.
This is why we ask you, the user, to define what projections you are
interested in: storing aggregates for the slices of the cube that interest you
is both a time and space save, and in many cases the only feasible solution.

@@

[What are the limits of our system? How high cardinality do we support? (This
is closely tied to hbase and the qualifier-limit.) If Martin is going to have
\textit{normalized\_path} as one of his dimensions, we might run into trouble.]

@@

Dumbo expects a key-value pair from the mapper. Hadoop will collect all keys
and give the reducer a key and an iterator of values. If we are to store
projections, we need to encode both the desired dimensions and the unit in
question in the key and emit the value of the unit in the \textit{value} field.
This way we can use a simple SumReducer to sum the values for each unit at any
point in n-space and be done.
